How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-17 tse-paper/content.tex
move chunk example to section about techniques

\begin{figure*}[!t]
  \centering
\begin{subfigure}[t]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* Try:
  \end{lstlisting}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[t]{\columnwidth}
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition,
RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}, escapeinside=//}
\begin{figure}[!t]
  \centering
  \lstinputlisting[breaklines=true]{listings/prose.txt}
  \caption{Simiplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example}}
  \label{lst:prose-program-simplified}
\end{figure}

\lstset{
  language=,
  morekeywords={},
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \lstinputlisting[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]{listings/chunk2.txt}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk showing a linter error}
  \label{lst:chunk-example-3}
\end{figure}

% TODO: this could move to the chunk retrieval techniques section
The results of our empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example} shows two exemplary
log chunks from Android build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed.
When PBE is trained with these two examples it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators, consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
cruicial for regular expressions to identify a log chunk.
In fact, if the third example from \Cref{lst:chunk-example-3} with a
different structural representation is added to the training set,
PBE is not able to syntesize a program and returns
``\texttt{no program found}''.
We express this difference in structural representation through
dividing log chunks into structural categories.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a \emph{different structural category}.

@@

@@

\subsection{Study Process}
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---end therefore the
same build environment (1).
One to five examples make up the training set (2).
We vary its size to measure how many examples are needed to
confidently configure a technique.
The training set is so small, because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort.
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a training set size of 1, as users of our techniques
should receive reliable results on every build log they analyze.
As a next step, we configure the chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
\begin{itemize}
  \item The size of training set (\textbf{RQ2.1}),
  \item the number of structural categories in the training set
  (\textbf{RQ2.2})
  \item the lines of the retrieval output ($\mathit{RetrievedLines}$),
  and our oracle, the output defined in the test example
  ($\mathit{DesiredLines}$).
  From these we calculate accuracy metrics (\textbf{RQ2.1}).
\end{itemize}

@@

\textbf{Refinement of Retrieval Quality for each Technique} We
investigated basic configurations of existing techniques applied to
chunk retrieval from build logs.
In a next step, each of these
techniques could be refined to better approach the domain of build
logs.
The \emph{LogChunks} data set and our study results act as a
baseline to benchmark such technique improvements.
We propose the
following improvements:
\begin{itemize}
  \item \textbf{Custom Ranking and Tokens for PBE} The program
  synthesis through PROSE ranks possible programs according to
  what the user most likely intended.
  One could adapt the ranking
  rules provided by the FlashExtract DSL to fit common build log
  chunk retrieval tasks.
  FlashExtract includes special tokens when
  enumerating possible regular expressions.
  One could extend these
  with tokens found in build logs, such as ``-'',``='',``ERROR''
  or ``[OK''.
  \item \textbf{Meta-Parameter Optimization for CTS} Information
  retrieval techniques have various meta-parameters which can be
  optimized for the specific use
  case~\cite{panichella2016parameterizing}.
  We propose to further
  investigate improvements in preprocessing of the log text, in
  tokenization of the log lines into terms and in stop words
  lists.
\end{itemize}

>>>>

@@

\subsection{Chunk Retrieval Example}
\begin{figure*}[!htbp]
  \centering
\begin{subfigure}[!htbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* Try:
  \end{lstlisting}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[!htbp]{\columnwidth}
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition,
RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}, escapeinside=//}
\begin{figure}[!htbp]
  \centering
  \lstinputlisting[breaklines=true]{listings/prose.txt}
  \caption{Simplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example}}
  \label{lst:prose-program-simplified}
\end{figure}

\lstset{
  language=,
  morekeywords={},
  texcl=false
}
\begin{figure}[!htbp]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \lstinputlisting[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]{listings/chunk2.txt}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk showing a linter error}
  \label{lst:chunk-example-3}
\end{figure}

To illustrate the chunk retrieval techniques further, we want to give
an example of real-world log chunks and how the techniques can
retrieve them from build logs.
\Cref{lst:chunk-example} shows two exemplary
log chunks from Android build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed, the information most often targeted by
other articles which analyze build logs.

When PBE is trained with these two examples it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators, consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if the third example from \Cref{lst:chunk-example-3} with a
different structural representation is added to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}''.
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of our empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

CTS will rank the lines from the analyzed build log which are most
similar to the chunks in the training examples,
marked blue in \Cref{lst:chunk-example}.
It will then return the two lines ranked highest, as the chunks in
the training examples have on average two lines.

KWS is configured with keywords that appear close to the chunk within
the build log.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE'', ``wrong'' or ``failed''.
The example from \Cref{lst:chunk-example-3} could be found with
``FAIL''.
If KWS would be trained with all three examples and the afformentioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice, most often within
the training set of three examples.

@@

\subsection{Study Process}
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---end therefore the
same build environment (1).
One to five examples make up the training set (2).
We vary its size to measure how many examples are needed to
confidently configure a technique.
The training set is so small, because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort.
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a training set size of 1, as users of our techniques
should receive reliable results on every build log they analyze.
As a next step, we configure the chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
  \item The size of training set (\textbf{RQ2.1}),
  \item the number of structural categories in the training set
  (\textbf{RQ2.2})
  \item the lines of the retrieval output ($\mathit{RetrievedLines}$),
  and our oracle, the output defined in the test example
  ($\mathit{DesiredLines}$).
  From these we calculate accuracy metrics (\textbf{RQ2.1}).
\end{itemize}

@@

\textbf{Refinement of Retrieval Quality for each Technique} We
investigated basic configurations of existing techniques applied to
chunk retrieval from build logs.
In a next step, each of these
techniques could be refined to better approach the domain of build
logs.
The \emph{LogChunks} data set and our study results act as a
baseline to benchmark such technique improvements.
We propose the
following improvements:
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
  \item \textbf{Custom Ranking and Tokens for PBE} The program
  synthesis through PROSE ranks possible programs according to
  what the user most likely intended.
  One could adapt the ranking
  rules provided by the FlashExtract DSL to fit common build log
  chunk retrieval tasks.
  FlashExtract includes special tokens when
  enumerating possible regular expressions.
  One could extend these
  with tokens found in build logs, such as ``-'',``='',``ERROR''
  or ``[OK''.
  \item \textbf{Meta-Parameter Optimization for CTS} Information
  retrieval techniques have various meta-parameters which can be
  optimized for the specific use
  case~\cite{panichella2016parameterizing}.
  We propose to further
  investigate improvements in preprocessing of the log text, in
  tokenization of the log lines into terms and in stop words
  lists.
\end{itemize}
